---
title: "Gut microbiota features of the geographically diverse Indian population"
author: "Sudarshan Shetty"
date: "November 12, 2018"
output: html_document
---

# Load libraries  
```{r, message=FALSE, warning=FALSE}

library(microbiome)
library(phyloseq)
library(microbiomeutilities)
library(ggpubr)
library(RColorBrewer)
library(vegan)

```



```{r}

#ps0 <-read_phyloseq(otu.file = "./pseq_input/pan_otutab.csv",
#                    taxonomy.file = "./pseq_input/pan_taxonomy.csv",
#                    metadata.file = "./pseq_input/pan_meta.csv",
#                    type = "simple")

#unique(tax_table(ps0)[,1])

```


```{r}
#ps0 <- remove_samples("Subject-8032",ps0)
#ps0
#rank_names(ps0)

```


```{r}

#saveRDS(ps0, "ps0.rds")
#rich <- plot_richness()
ps0 <- readRDS("ps0.rds")

head(meta(ps0))

ps0.h <- subset_samples(ps0, Obese.Non.Obese != "Obese")
rank_names(ps0.h)
```



# Figure 1 

Phylum Level:  
Figure 1: Comparison of relative abundances of major phyla in the gut microbiota of Indians. The p-values were calculated using Wilcoxon test.  

```{r}

ps0.phy <- aggregate_taxa(ps0, "Phylum", top = 6)

metadf1 <- meta(ps0.phy)
# create comaprision
lev1 <- levels(metadf1$Geographical.zone.in.India) # get the variables

# make a pairwise list that we want to compare.
L.pairs1 <- combn(seq_along(lev1), 2, simplify = FALSE, FUN = function(i)lev1[i])

ps0_ldf <- phy_to_ldf(ps0.phy, transform.counts =  NULL)

ps0_ldf.sub <- subset(ps0_ldf, OTU != "Other")
ps0_ldf.sub2 <- subset(ps0_ldf.sub, Obese.Non.Obese != "Obese")

px2 <- ggboxplot(ps0_ldf.sub2, x = "Geographical.zone.in.India", y = "Abundance", 
                 #add = "dotplot", 
                 fill = "Geographical.zone.in.India", 
                 palette = c("#e41a1c", "#377eb8", "#4daf4a","#984ea3"), 
                 facet.by = "OTU", 
                 legend = "NULL",
                 ylab = "Relative abundance (Log10)") + stat_compare_means(comparisons = L.pairs1) + scale_y_log10() + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))
px2
ggsave("./figures/figure 1 top phyla.pdf", height = 10, width = 12)
ggsave("./figures/figure 1 top phyla.png", height = 10, width = 12)

```


# Figure 2  

Figure 2: Inter-individual variation in relative abundance of top 10 gut microbial genera in subject from different geographical zones in India.  

```{r}

#tab <- inequality(ps0)

taxic <- as.data.frame(ps0@tax_table)  # this will help in setting large color options


taxic$OTU <- rownames(taxic)  # Add the OTU ids from OTU table into the taxa table at the end.
colnames(taxic)  # Y

ps0.h.lev <- ps0

taxmat <- as.matrix(taxic)  # convert it into a matrix.
new.tax <- tax_table(taxmat)  # convert into phyloseq compatible file.
tax_table(ps0.h.lev) <- new.tax  # incroporate into phyloseq Object

ps0.h.levf <- format_phyloseq(ps0.h.lev)

colnames(tax_table(ps0.h.levf)) <- c("Phylum",  "Class",   "Order",   "Family",  "Genus",   "Species", "OTU")

ps1.com.fam <- microbiome::aggregate_taxa(ps0.h.levf, "Genus", top = 25)

#colourCount = length(unique(taxic$Family))  #define number of variable colors based on number of Family (change the level accordingly to phylum/class/order)
#getPalette = colorRampPalette(brewer.pal(12, "Paired"))  # change the palette as well as the number of colors will change according to palette.


guide_italics <- guides(fill = guide_legend(label.theme = element_text(size = 15, 
    face = "italic", colour = "Black", angle = 0)))

plot.composition.relAbun <- plot_composition(ps1.com.fam, 
                                             sample.sort = "Geographical.Location", 
                                             x.label = "Geographical.zone.in.India") + theme(legend.position = "bottom") + scale_fill_brewer("Genus", palette = "Paired") + theme_bw() + 
    theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("Relative abundance") + guide_italics + theme(legend.title = element_text(size=18))
  

data.com <- plot.composition.relAbun$data
colnames(data.com)
p.heat <- ggplot(data.com, aes(x = Sample, y = OTU)) + geom_tile(aes(fill = Abundance)) 

# Change color
p.heat <- p.heat + scale_fill_distiller("Abundance", palette = "RdYlBu") + theme_bw() 

# Make bacterial names italics
p.heat <- p.heat + theme(axis.text.y = element_text(colour = 'black', 
                                                    size = 10, 
                                                    face = 'italic')) 
# Make seperate samples based on main varaible
p.heat <- p.heat + facet_grid(~xlabel, 
                              scales = "free") + rremove("x.text") 

p.heat <- p.heat + ylab("Taxa")

#Clean the x-axis
p.heat <- p.heat + theme(axis.title.x=element_blank(),
                     axis.text.x=element_blank(),
                     axis.ticks.x=element_blank()) 

# Clean the facet label box
p.heat <- p.heat + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))

print(p.heat)

#dir.create("figures")

ggsave("./figures/Figure 2 Genus_heatmap.pdf", height = 6, width = 8)
ggsave("./figures/Figure 2 Genus_heatmap.tiff", height = 6, width = 8)
ggsave("./figures/Figure 2 Genus_heatmap.png", height = 6, width = 8)

```


# Figure 3  

Figure 3: Principal coordinates analysis (PCoA) based on Bray-Curtis dissimilarity based on OTU relative abundances.   
```{r}

ps0.f <- format_to_besthit(ps0)

pan.ord.mds.otu <- ordinate(ps0.f, "MDS", "bray")

myPalette <- colorRampPalette(rev(brewer.pal(11, "Set1")))

sc.man <- scale_colour_manual(values = myPalette(14))

p_pan_otu <- plot_ordination(ps0.f, pan.ord.mds.otu, type = "samples", color = "Geographical.Location", shape = "Geographical.zone.in.India")

p_pan_otu <- p_pan_otu + sc.man + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) #+ theme(legend.position="bottom")
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
p_pan_otu

ggsave("./figures/Figure 3 pan_otu_all.pdf", height = 6, width = 10)
ggsave("./figures/Figure 3 pan_otu_all.tiff", height = 6, width = 10)
ggsave("./figures/Figure 3 pan_otu_all.png", height = 6, width = 10)


```

# permanova  

```{r}

summary(taxa_sums(ps0))
ps0.fil <- core(ps0, 0.00001, 0.05)
ps0.ord.mds <- metaMDS(t(abundances(ps0.fil)), k=3)

#ps0.ord.mds <- ordinate(ps0, "MDS", "bray")

fit <- envfit(ps0.ord.mds, meta(ps0.fil), perm = 999)

scores(fit, "vectors")
plot(ps0.ord.mds)
plot.envfit(fit)

set.seed(9795769)

ps0.otu <- abundances(ps0.f)
ps0.metadf <- meta(ps0.f)
permanova.obnob <- adonis(t(ps0.otu) ~ Obese.Non.Obese,
               data = ps0.metadf, permutations=999, method = "bray")
permanova.obnob
# P-value
print(as.data.frame(permanova.obnob$aov.tab)["Obese.Non.Obese", "Pr(>F)"])
dist <- vegdist(t(ps0.otu))
anova(betadisper(dist, ps0.metadf$Obese.Non.Obese))
# 0.57
# Geographical.Location
# Geographical.zone.in.India
# Gender
# AGE..in.years
# Life.style.pattern
# BMI
permanova.city <- adonis(t(ps0.otu) ~ Geographical.Location,
               data = ps0.metadf, permutations=999, method = "bray")

# P-value
permanova.city

coef <- coefficients(permanova.city)["Geographical.Location2",]
top.coef <- coef[rev(order(abs(coef)))[1:20]]
par(mar = c(3, 14, 2, 1))
barplot(sort(top.coef), horiz = T, las = 1, main = "Top taxa")



permanova.zone <- adonis(t(ps0.otu) ~ Geographical.zone.in.India,
               data = ps0.metadf, permutations=999, method = "bray")

# P-value
permanova.zone


permanova.gener <- adonis(t(ps0.otu) ~ Gender,
               data = ps0.metadf, permutations=999, method = "bray")

# P-value
permanova.gener


permanova.Life.style.pattern <- adonis(t(ps0.otu) ~ Life.style.pattern,
               data = ps0.metadf, permutations=999, method = "bray")

# P-value
permanova.Life.style.pattern
```

# perdisp 

```{r}

bray.dist <- phyloseq::distance(ps0.occ, "bray")

ps0.otu <- abundances(ps0.occ)
ps0.metadf <- meta(ps0.occ)


Obese.Non.Obese.disp <- anova(betadisper(bray.dist, ps0.metadf$Obese.Non.Obese))
Obese.Non.Obese.disp 
boxplot(betadisper(bray.dist, ps0.metadf$Obese.Non.Obese))

Geographical.Location.disp <- anova(betadisper(bray.dist, ps0.metadf$Geographical.Location))
Geographical.Location.disp
boxplot(betadisper(bray.dist, ps0.metadf$Geographical.Location), las = 2)
plot(betadisper(bray.dist, ps0.metadf$Geographical.Location))

Geographical.zone.in.India.disp <- anova(betadisper(bray.dist, ps0.metadf$Geographical.zone.in.India))
Geographical.zone.in.India.disp

Gender.disp <- anova(betadisper(bray.dist, ps0.metadf$Gender))
Gender.disp

Life.style.pattern.disp <- anova(betadisper(bray.dist, ps0.metadf$Life.style.pattern))
Life.style.pattern.disp



```

#Figure 4  

Figure 4: Occupancy-Abundance relationship for OTUs from major phyla in the Indian gut microbiota (n=1003). The x-axis is log transformed for clarity.  

Occupancy test  

```{r}

#p <- plot_taxa_prevalence(ps0.h.lev, 'Phylum', detection = 0)
#print(p)

```



```{r}

ps0.occ <- format_to_besthit(ps0)
colnames(tax_table(ps0.occ)) <- c("Phylum", "Class", "Order", "Family", "Genus", "Species", "OTUID","best_hit")
otu.abun = apply(otu_table(ps0.occ),1,mean)
head(tax_table(ps0))
# Calculate the frequency of each OTU across all samples
# 
otu.freq = rowSums(otu_table(ps0.occ) != 0)/nsamples(ps0.occ)

# Reassign names of phyla so we only color by the top 5 phyla and mark all others as "other"
phyla = as.vector(data.frame(tax_table(ps0.occ))$Phylum)

levels(phyla) = c(levels(phyla),"other")
#
keephyla = c("Actinobacteria","Bacteroidetes","Firmicutes","Proteobacteria","Verrucomicrobia", "Spirochaetes")

phyla[!(phyla %in% keephyla)] = "Other"

phyla = as.vector(phyla)

phyla=as.factor(phyla)


otuabun = data.frame(abundance=log(otu.abun),frequency=otu.freq,phyla)
#fo saving rel abun
otuabun2 = data.frame(abundance=otu.abun,frequency=otu.freq,phyla)
write.csv(otuabun2, "occup-abund.csv")

# Use color brewer to pick a color scheme for the phyla
brew = brewer.pal(7, "Paired")

# Create a scatterplot of OTUs showing their average relative abundance and frequency 
# This plot shows how rare and abundant OTUs are distributed across the
# your study.

p.occ <- ggplot(otuabun, aes(x=abundance,y=frequency)) + geom_point(aes(size = frequency), shape = 1, alpha = 0.9) +
    xlab("Average relative abundance (log scale)") + 
    ylab("OTU Frequency") + facet_wrap(~phyla) +
    theme_bw() +
    theme(
        plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=10)
    ) + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))
	
p.occ

ggsave("./figures/Figure 4 Occupancy_all.pdf", height = 4, width = 8)
ggsave("./figures/Figure 4 Occupancy_all.tiff", height = 4, width = 8)
ggsave("./figures/Figure 4 Occupancy_all.png", height = 4, width = 8)
```



# Figure 6  

**Prevotell/Bacteroides**   

The codes for this analysis were modified from [Gorvitovskaia et al., 2016](https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-016-0160-7)  

Figure 6: Principal coordinates analysis based on Bray-Curtis dissimilarity based on genus level relative abundances. A] PCoA depicting the gradient of Prevotella/Bacteroides ratio and distribution of body mass index (BMI) in 1003 Indian subjects. B] Same PCoA as in panel A, but coloured and facetted for depicting the distribution of Prevotella/Bacteroides ratio in the Indian gut microbiota in different geographical zones (East, n = 250; North, n = 243; South, n = 250; West, n = 260).  

```{r}

ps0.fv <- format_phyloseq(ps0)
head(tax_table(ps0.fv))

colnames(tax_table(ps0.fv)) <- c("Phylum",  "Class",   "Order",   "Family",  "Genus",   "Species", "OTU")

ps0.gen.all <- microbiome::aggregate_taxa(ps0.fv, "Genus")

relab2 <- abundances(ps0.gen.all)

#Finding the Prevotella dominated samples in a given matrix
findPrevoDom <- function(theMatrix) {
  y <- apply(theMatrix, 2, function(x) {order(x, decreasing=TRUE)[1] == grep("Prevotella", rownames(theMatrix), perl=TRUE)})
  return(theMatrix[,y])
}

#Finding the Bacteroides dominated samples in a given matrix
findBacteDom <- function(theMatrix) {
  y <- apply(theMatrix, 2, function(x) {order(x, decreasing=TRUE)[1] == grep("Bacteroides", rownames(theMatrix), perl=TRUE)})
  return(theMatrix[,y])
}

pDom <- colnames(findPrevoDom(relab2))
bDom <- colnames(findBacteDom(relab2))
#oDom <- colnames(otumat)[ -which(colnames(otumat) %in% c(pDom,bDom))]
dominance <- rep("Other",1003)
names(dominance) <- sample_names(ps0.gen.all)
dominance[pDom] <- "Prevotella Dominance"
dominance[bDom] <- "Bacteroides Dominance"
# To check
dominance[1:30]
pDom[1:30]
bDom[1:8]

#rownames(tax_table(ps0.gen.all))[tax_table(ps0.gen.all)[,1]=="Bacteroides"]
#rownames(tax_table(ps0.gen.all))[tax_table(ps0.gen.all)[,1]=="Prevotella"]

# otu58 bacteroides
# otu265 prev


bact_otu <- "Bacteroides"
prev_otu <- "Prevotella"

#tax_table(ps0.gen.all)["otu58",]
#tax_table(ps0.gen.all)["otu265",]

pb_ratio <- rep(0,1003)
names(pb_ratio) <- sample_names(ps0.gen.all)
pb_ratio <- apply(otu_table(ps0.gen.all),2,function(x) {x[prev_otu]/(x[prev_otu] + x[bact_otu])})

pb_binary <- rep(0,1003)
pb_binary[pb_ratio>0.5] = 1

bact_abun <- rep(0,1003)
names(bact_abun) <- sample_names(ps0.gen.all)
bact_abun <- apply(otu_table(ps0.gen.all),2,function(x) {x[bact_otu]})

prev_abun <- rep(0,1003)
names(prev_abun) <- colnames(otu_table(ps0.gen.all))
prev_abun <- apply(otu_table(ps0.gen.all),2,function(x) {x[prev_otu]})



#Just checking
otu_table(ps0.gen.all)[prev_otu,4]/(otu_table(ps0.gen.all)[prev_otu,4] + otu_table(ps0.gen.all)[bact_otu,4])
pb_ratio[4]

otu_table(ps0.gen.all)[bact_otu,26]
bact_abun[26]
otu_table(ps0.gen.all)[prev_otu,26]
prev_abun[26]



sampledata = sample_data(data.frame(dominance=dominance, 
                                    pb_ratio=pb_ratio, bact_abun=bact_abun, 
                                    prev_abun=prev_abun, pb_binary=pb_binary, 
                                    row.names = sample_names(ps0.gen.all), 
                                    stringsAsFactors = FALSE))
sampledata[1:5,]
sample_data(ps0.gen.all)$dominance <- factor(dominance)

ps0.gen.all3 = merge_phyloseq(ps0.gen.all, sampledata)
ps0.gen.all3
meta(ps0.gen.all3)

table(meta(ps0.gen.all3)$Geographical.zone.in.India)
table(meta(ps0.gen.all3)$Gender)

saveRDS(ps0.gen.all3, "ps0.gen.all3.rds")
ps0.gen.all3 <- readRDS("ps0.gen.all3.rds")
```

# ordination 2
```{r}

ob.ord.mds <- ordinate(ps0.gen.all3, "MDS", "bray")

###Dominance Plots
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
sc <- scale_colour_gradientn(colours = myPalette(100))
p_ob_pb <- plot_ordination(ps0.gen.all3, ob.ord.mds, type = "samples", color = "pb_ratio")
p_ob_pb <- p_ob_pb + geom_point(aes(size = BMI)) + sc + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) 
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
p_ob_pb

ggsave("./figures/Ord_pb_ratio_obse_nob.pdf", height = 4, width = 6)
ggsave("./figures/Ord_pb_ratio_obse_nob.tiff", height = 4, width = 6)
ggsave("./figures/Ord_pb_ratio_obse_nob.png", height = 4, width = 6)

p_ob_pb2 <- plot_ordination(ps0.gen.all3, ob.ord.mds, type = "samples", color = "pb_ratio", shape = "Geographical.zone.in.India")
p_ob_pb2 <- p_ob_pb2 + sc + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) 
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
p_ob_pb2
ggsave("./figures/Ord_pb_ratio_zone_nob.pdf", height = 4, width = 6)
ggsave("./figures/Ord_pb_ratio_zone_nob.tiff", height = 4, width = 6)
ggsave("./figures/Ord_pb_ratio_zone_nob.png", height = 4, width = 6)
```


```{r}

p_ob_pb2 <- plot_ordination(ps0.gen.all3, ob.ord.mds, type = "samples", color = "pb_ratio")
p_ob_pb2 <- p_ob_pb2 + sc + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
p_ob_pb2 <-p_ob_pb2 + facet_wrap(~Geographical.zone.in.India, scales="free")
```


```{r}

pala <- ggarrange(p_ob_pb, p_ob_pb2, nrow = 2, labels = c("A", "B"))
pala
ggsave("./figures/Figure 6 ord_all.pdf", height = 8, width = 8)
ggsave("./figures/Figure 6 ord_all.tiff", height = 8, width = 8)
ggsave("./figures/Figure 6 ord_all.png", height = 8, width = 8)
```

# Supplementary figure 5   

**Corelation between axis 1 and P/B ratio**

Supplementary figure 5: Pearson’s correlation between PCoA axis 1 and P/B ratio. 
```{r}

p.ob_axis <- p_ob_pb2$data

pc.pb2 <- ggscatter(p.ob_axis, "Axis.1", "pb_ratio",  add = "reg.line",
                    facet.by = "Geographical.zone.in.India", 
                    shape = 21, size = 3,
                    conf.int = TRUE,
                    xlab = "PCoA 1",
                    ylab = "P/B ratio")

pc.pb2 <- pc.pb2 + stat_cor(method = "pearson", label.x = -0.3, label.y = 0.25)

pc.pb2 <- pc.pb2 + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))

pc.pb2
ggsave("./figures/PCoA1_PB_ratio.pdf", height = 8, width = 8)
ggsave("./figures/PCoA1_PB_ratio.tiff", height = 8, width = 8)
ggsave("./figures/PCoA1_PB_ratio.png", height = 8, width = 8)
```




# Regionwise  
## Supplementary Figure 1  
## South

```{r}

ps0.gen.all3.south <- subset_samples(ps0.gen.all3, Geographical.zone.in.India == "South")
ps0.gen.all3.south <- prune_taxa(taxa_sums(ps0.gen.all3.south) > 0, ps0.gen.all3.south)
ps0.gen.all3.south.mds.ord <- ordinate(ps0.gen.all3.south, "MDS", "bray")

permanova.south <- adonis(t(abundances(ps0.gen.all3.south)) ~ Geographical.Location,
               data = meta(ps0.gen.all3.south), permutations=999, method = "bray")

myPalette <- colorRampPalette(rev(brewer.pal(11, "Set1")))

sc.man <- scale_colour_manual(values = myPalette(14))

south.mds.ord <- plot_ordination(ps0.gen.all3.south, ps0.gen.all3.south.mds.ord, 
                             type = "samples", color = "Geographical.Location" 
                             ) + geom_point(size = 3)

south.mds.ord <- south.mds.ord + sc.man + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) #+ theme(legend.position="bottom")
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
south.mds.ord + stat_ellipse()

```

## North  

```{r}
ps0.gen.all3.North <- subset_samples(ps0.gen.all3, Geographical.zone.in.India == "North")
ps0.gen.all3.North <- prune_taxa(taxa_sums(ps0.gen.all3.North) > 0, ps0.gen.all3.North)
ps0.gen.all3.North.mds.ord <- ordinate(ps0.gen.all3.North, "MDS", "bray")

permanova.north <- adonis(t(abundances(ps0.gen.all3.North)) ~ Geographical.Location,
               data = meta(ps0.gen.all3.North), permutations=999, method = "bray")
permanova.north
#myPalette <- colorRampPalette(rev(brewer.pal(6, "Paired")))

#sc.man <- scale_colour_manual(values = myPalette(14))

north.mds.ord <- plot_ordination(ps0.gen.all3.North, ps0.gen.all3.North.mds.ord, 
                             type = "samples", color = "Geographical.Location" 
                             ) + geom_point(size = 3) + scale_colour_brewer(palette = "Set2")

north.mds.ord <- north.mds.ord + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) #+ theme(legend.position="bottom")
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
north.mds.ord + stat_ellipse()
```

## West  

```{r}
ps0.gen.all3.west <- subset_samples(ps0.gen.all3, Geographical.zone.in.India == "West")
ps0.gen.all3.west <- prune_taxa(taxa_sums(ps0.gen.all3.west) > 0, ps0.gen.all3.west)
ps0.gen.all3.west.mds.ord <- ordinate(ps0.gen.all3.west, "MDS", "bray")

permanova.west <- adonis(t(abundances(ps0.gen.all3.west)) ~ Geographical.Location,
               data = meta(ps0.gen.all3.west), permutations=999, method = "bray")
permanova.west

#myPalette <- colorRampPalette(rev(brewer.pal(6, "Paired")))

#sc.man <- scale_colour_manual(values = myPalette(14))

west.mds.ord <- plot_ordination(ps0.gen.all3.west, ps0.gen.all3.west.mds.ord, 
                             type = "samples", color = "Geographical.Location" 
                             ) + geom_point(size = 3) + scale_colour_brewer(palette = "Set3")

west.mds.ord <- west.mds.ord + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) #+ theme(legend.position="bottom")
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
west.mds.ord + stat_ellipse()
```


## East  

```{r}
ps0.gen.all3.east <- subset_samples(ps0.gen.all3, Geographical.zone.in.India == "East")
ps0.gen.all3.east <- prune_taxa(taxa_sums(ps0.gen.all3.east) > 0, ps0.gen.all3.east)
ps0.gen.all3.east.mds.ord <- ordinate(ps0.gen.all3.east, "MDS", "bray")

permanova.east <- adonis(t(abundances(ps0.gen.all3.east)) ~ Geographical.Location,
               data = meta(ps0.gen.all3.east), permutations=999, method = "bray")
permanova.east
permanova.east$coefficients
#myPalette <- colorRampPalette(rev(brewer.pal(6, "Paired")))

#sc.man <- scale_colour_manual(values = myPalette(14))

east.mds.ord <- plot_ordination(ps0.gen.all3.east, ps0.gen.all3.east.mds.ord, 
                             type = "samples", color = "Geographical.Location"
                             ) + geom_point(size = 3) + scale_colour_manual(values = rev(brewer.pal(9, "Paired")))

east.mds.ord <- east.mds.ord + theme_bw() +
  theme(plot.background = element_blank()
        ,panel.grid.major = element_blank()
        ,panel.grid.minor = element_blank()
        ,panel.background = element_blank()
        ,axis.text.x  = element_text(angle=90, vjust=0.5, size=12)
    ) #+ theme(legend.position="bottom")
	
  #annotate("text",label="(B)",x = -0.5, y =0.29 , size=7)
east.mds.ord$data

#east.mds.ord + stat_ellipse(geom = "polygon", type="norm", alpha=0.4, aes(fill=Geographical.Location))
```

```{r}

p.zone <- ggarrange(east.mds.ord, west.mds.ord, north.mds.ord, south.mds.ord, 
                    ncol = 2, nrow = 2, labels = c("A", "B", "C", "D"))
p.zone
ggsave("./figures/Supplementary Figure S1 zone_indiv_pcoa.pdf", height = 6, width = 10)
ggsave("./figures/Supplementary Figure S1 zone_indiv_pcoa.tiff", height = 6, width = 10)
ggsave("./figures/Supplementary Figure S1 zone_indiv_pcoa.png", height = 6, width = 10)
```

## Supplementary figure 3  
```{r}

ps0.gen.all3 <- readRDS("ps0.gen.all3.rds")
write_phyloseq(ps0.gen.all3, "OTU")
meta_gendf <- meta(ps0.gen.all3)
colnames(meta_gendf)
p.ag <- ggscatter(meta_gendf, "AGE..in.years", "pb_ratio",
                  color = "Obese.Non.Obese", xlab = "Age (years)", ylab = "P/B ratio" ,
                  palette = c("#e41a1c", "#377eb8", "#4daf4a","#984ea3")) + stat_cor(label.x = 40, label.y = 0.1)
p.ag 

p.bm <- ggscatter(meta_gendf, "BMI", "pb_ratio",
                  color = "Gender", xlab = "BMI", ylab = "P/B ratio" ,
                  palette = c("#4daf4a","#984ea3")) + stat_cor(label.x = 30, label.y = 0.1)
p.bm

p.ag.bm <- ggarrange(p.ag, p.bm, labels = c("A", "B"))
ggsave("./figures/Supplementary figure 3 scatter_all.pdf", height = 4, width = 8)
ggsave("./figures/Supplementary figure 3 scatter_all.tiff", height = 4, width = 8)
ggsave("./figures/Supplementary figure 3 scatter_all.png", height = 4, width = 8)
```


# Core microbiota  

```{r}

# ps0.occ <- format_to_besthit(ps0)
#saveRDS(ps0.occ, "ps0.occ.rds")
ps0.occ <- readRDS("ps0.occ.rds")
det <- c(0, 0.001, 0.01, 0.1, 0.5, 2, 5, 20)/100
preval <- seq(.05, 1, .05)

p.core.lne <- plot_core(ps0.occ, prevalences = preval, detections = det, plot.type = "lineplot") + xlab("Relative Abundance (%)") + scale_color_distiller(palette = "Spectral")
p.core.lne

ggsave("./figures/lne Pan India Core Microbiota.pdf", height = 4, width = 6)
ggsave("./figures/lne Pan India Core Microbiota.png", height = 4, width = 6)
prevalences <- seq(.05, 1, .05)
detections <- 10^seq(log10(1e-4), log10(.2), length = 10)

pancore <- plot_core(ps0.occ, plot.type = "heatmap", 
                         prevalences = prevalences,
                         detections = detections,
                         colours = rev(brewer.pal(5, "RdYlBu")),
                         min.prevalence = .75, horizontal = F)


# sel
pancore <- pancore + theme_bw(base_size = 10)+ theme(axis.text.y = element_text(face="italic")) 

pancore <- pancore + ggtitle("Pan India Core Microbiota (1003 healthy subjects)")

pancore

core_data <- pancore$data
write.table(core_data, "Pan_Core_det_prev_values.txt", sep="\t")
ggsave("./figures/Pan India Core Microbiota.pdf", height = 6, width = 7)
ggsave("./figures/Pan India Core Microbiota.png", height = 6, width = 7)

```

## Identify core Bootstrap OTU-level

```{r}
source("./core_scripts/bootstrap_core_estimation.R")

otu.tb <- abundances(ps0.occ)

cor.boot <- bootstrap.microbes(otu.tb, Nsample = NULL, minPrev = 810, Nboot = 1000, 
                               I.thr = 0.0001, ncore = 1, I.max=NULL)
cor.boot$Microbe
write.csv(cor.boot, "cor.boot.csv")

DT::datatable(cor.boot)
```



## Identify core Bootstrap Genus-level  
```{r}
#source("./core_scripts/bootstrap_core_estimation.R")

otu.gen.tb <- abundances(ps0.gen.all3)

cor.gen.boot <- bootstrap.microbes(otu.gen.tb, Nsample = NULL, minPrev = 810, Nboot = 1000, 
                               I.thr = 0.0001, ncore = 1, I.max=NULL)
#cor.gen.boot$Microbe
write.csv(cor.gen.boot, "cor.gen.boot.csv")

DT::datatable(cor.gen.boot)
```


## Bootstrap Core abundance  

## Supplementary figure 2  

```{r}

core.gen <- subset(cor.gen.boot, suggested.core == 1)
core.gne.f <- as.character(core.gen$Microbe)
core.ps0.gen <- prune_taxa(core.gne.f, ps0.gen.all3)

core.gen.ab_0.001 <- sample_sums(core.ps0.gen)

sample_data(core.ps0.gen)$coreAb_0.001 <- core.gen.ab_0.001
coreAB.gen <- ggboxplot(meta(core.ps0.gen), "Gender", "coreAb_0.001", 
                   #color = "BMI", 
                   facet.by = "Geographical.zone.in.India",
                   add= "jitter", 
                   xlab = "Geographical zone in India",
                   ylab = "Core genus abundance\n(Detection = 0.0001 and Prevalence = 80%)")
coreAB.gen <- coreAB.gen + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))
coreAB.gen
ggsave("./figures/Supplementary figure 2 Pan India Core genus abudnace.pdf", height = 6, width = 7)
ggsave("./figures/Supplementary figure 2 Pan India Core genus abundance.png", height = 6, width = 7)
```

**OTU-lev**  

```{r}

core.taxa.standard <- core_members(ps0.occ, detection = .0001, prevalence = .95)

core.abundance_0.001 <- sample_sums(core(ps0.occ, detection = .0001, prevalence = .95))

sample_data(ps0.occ)$coreAb_0.001 <- core.abundance_0.001

ps0.occ_meta <- meta(ps0.occ)
colnames(ps0.occ_meta)
coreAB <- ggboxplot(ps0.occ_meta, "Gender", "coreAb_0.001", 
                   #color = "BMI", 
                   facet.by = "Geographical.zone.in.India",
                   add= "jitter", 
                   xlab = "Geographical zone in India",
                   ylab = "Core abundance\n(Detection = 0.0001 and Prevalence = 95%)")
coreAB <- coreAB + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))
coreAB

ggsave("./figures/Pan India Core abudnace.pdf", height = 6, width = 7)
ggsave("./figures/Pan India Core abundance.png", height = 6, width = 7)
```


## Supplementary figure 4  

```{r}
head(meta(ps0.gen.all3))

meta.pball <- meta(ps0.gen.all3)

meta.pball$logPBratio <- log10(meta.pball$pb_ratio)

# no dif gender
p.com <- ggboxplot(meta.pball, "Gender", "pb_ratio",
                   ylab = "P/B ratio",
                   facet.by = "Geographical.zone.in.India") 
p.com <- p.com + stat_compare_means() + theme(legend.key = element_blank(), 
                     strip.background = element_rect(colour="black", fill="white"))

p.com 

ggsave("./figures/pb_ratio_gender.pdf", height = 4, width = 6)
ggsave("./figures/pb_ratio_gender.tiff", height = 4, width = 6)
ggsave("./figures/pb_ratio_gender.png", height = 4, width = 6)


gloc <- levels(meta.pball$Geographical.zone.in.India) # get the variables

# make a pairwise list that we want to compare.
gloc.pairs <- combn(seq_along(gloc), 2, simplify = FALSE, FUN = function(i)gloc[i])

p.com2 <- ggboxplot(meta.pball, "Geographical.zone.in.India", "pb_ratio", facet.by = "Obese.Non.Obese") 
p.com2 <- p.com2 + stat_compare_means(comparisons = gloc.pairs) 
p.com2
str(meta.pball)

ggsave("./figures/pb_ratio_obse_nob.pdf", height = 4, width = 6)
ggsave("./figures/pb_ratio_obse_nob.tiff", height = 4, width = 6)
ggsave("./figures/pb_ratio_obse_nob.png", height = 4, width = 6)



ggscatter(meta.pball, "BMI", "pb_ratio", facet.by = "Geographical.zone.in.India", color = "Obese.Non.Obese") + stat_cor()

```


# Genus level differnces between metadata  

```{r}
ps0.gen.all3 <- readRDS("ps0.gen.all3.rds")

#dir.create("Stat_Tables")

# Base inputs  
source("./core_scripts/find_sigDiff_taxa.R")

# otutable
ts_fp <- as.data.frame.matrix(abundances(ps0.gen.all3)) 
# metadata
map_fp <- meta(ps0.gen.all3)

# Geographical zones 
taxa_diff.zone <- differences_in_taxa(ts_fp,
                                 map_fp,
                                 factor = 'Geographical.zone.in.India',
                                 filterLevel = 0.005,
                                 testType = 'KW'
                                 )

taxa_diff.zone.sig <- subset(taxa_diff.zone, pvalsBon <= 0.01)
head(taxa_diff.zone.sig)
write.csv(taxa_diff.zone.sig, "./Stat_Tables/Geographical.zone.in.India.csv")

# Geographical cities 
taxa_diff.city <- differences_in_taxa(ts_fp,
                                 map_fp,
                                 factor = 'Geographical.Location',
                                 filterLevel = 0.005,
                                 testType = 'KW'
                                 )

taxa_diff.city.sig <- subset(taxa_diff.city, pvalsBon <= 0.01)
head(taxa_diff.city.sig)
write.csv(taxa_diff.city.sig, "./Stat_Tables/Geographical.city.in.India.csv")


# Life.style.pattern 
taxa_diff.life <- differences_in_taxa(ts_fp,
                                 map_fp,
                                 factor = 'Life.style.pattern',
                                 filterLevel = 0.005,
                                 testType = 'MW'
                                 )

taxa_diff.life.sig <- subset(taxa_diff.life, pvalsBon <= 0.01)
head(taxa_diff.life.sig)
write.csv(taxa_diff.life.sig, "./Stat_Tables/LifeStylePattern.csv")

p.tax.Life.style.pattern <- plot_select_taxa2(ps0.gen.all3, rownames(taxa_diff.life.sig), variableA = "Life.style.pattern", "Paired", "violin")
p.tax.Life.style.pattern


# Obese 
taxa_diff.Obese.Non.Obese <- differences_in_taxa(ts_fp,
                                 map_fp,
                                 factor = 'Obese.Non.Obese',
                                 filterLevel = 0.005,
                                 testType = 'MW'
                                 )

taxa_diff.Obese.Non.Obese.sig <- subset(taxa_diff.Obese.Non.Obese, pvalsBon <= 0.01)
head(taxa_diff.Obese.Non.Obese.sig)
# No significant genus level

# Gender

taxa_diff.Gender <- differences_in_taxa(ts_fp,
                                 map_fp,
                                 factor = 'Gender',
                                 filterLevel = 0.005,
                                 testType = 'MW'
                                 )

taxa_diff.Gender.sig <- subset(taxa_diff.Gender, pvalsBon <= 0.01)
head(taxa_diff.Gender.sig)
# No significant genus level

taxa_diff.Gender.sig <- subset(taxa_diff.Gender, pvals <= 0.05)


```




